### 2013-09-25 nothing_to_hide/presentation.tex
Added initial shell.

\documentclass{beamer}
\usepackage{graphicx}

\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
\newcommand{\ts}{\textsuperscript}

\usetheme{Boadilla}

% items enclosed in square brackets are optional; explanation below
\title[Privacy in Art, Books, and Film]{I've Got Nothing to Hide: \\
  Exploring Privacy in Art, Books, and Film}
\author[Josh Datko]{Joshua Datko}
\institute[Drexel]{
  Department of Computer Science\\
  Drexel University\\
  Philadelphia, PA 19103\\[1ex]
  %\texttt{jbd65@drexel.edu}
}
\date[September 2013]{September 28, 2013}

\setbeamertemplate{bibliography item}[text]

\begin{document}

%--- the titlepage frame -------------------------%
\begin{frame}[plain]
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Motivation}

  \begin{itemize}
  \item Tor runs on volunteers.
  \item Tor needs more volunteers.
  \item Tor should incentivize relay and bridge operators.
  \end{itemize}

  \begin{block}{People like money.}
    Provide incentives for Tor server operators through Bitcoin.
  \end{block}
\end{frame}

% \begin{frame}
% \frametitle{Survival Curve}
%   \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{survival}
% \end{frame}
%% \begin{frame}
%%   \frametitle{Bitcoin}

%%   Bitcoin is a peer-to-peer electronic currency based on
%%   proof-of-work. \cite{nakamoto2009bitcoin}

%%   \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{bitcoin_volume}

%% \end{frame}

%% \begin{frame}
%%   \frametitle{Can Bitcoin provide incentives?}

%%   Bitcoin is not
%%   anonymous.\cite{journals/iacr/AndroulakiKRSC12} \cite{journals/corr/abs-1107-4524}

%%   Combining Bitcoin as is with Tor would expose Tor users.  We seek to
%%   investigate alternative means of payments.
%%   \begin{itemize}
%%   \item Tor client joins a Bitcoin mining pool and payment goes
%%     directly to the Tor Server?
%%   \item Can we use Zerocoin? \cite{Miers:2013:ZAD:2497621.2498124}
%%     \end{itemize}

%%   \begin{block}{Anonymity}
%%     Tor incentives can't compromise anonymity!!
%%   \end{block}
%% \end{frame}

%% \begin{frame}
%%   \frametitle{Approach}

%%   \begin{itemize}
%%   \item Compare BitTor to other incentive
%%     systems.\cite{Jansen:2010:RNT:1866307.1866344}\cite{lira-ndss13}\cite{raykova-pet2008}
%%   \item Investigate Bitcoin payment methods.
%%   \item Provide anonymity analysis.
%%     \item Prototype BitTor and investigate performance

%%     \end{itemize}

%%   \begin{block}{Live and dynamic systems}
%%     Bitcoin = money, Tor = privacy $\rightarrow$ testing to be done ``offline.''
%%   \end{block}
%% \end{frame}

%% \begin{frame}[allowframebreaks]

%%   \frametitle{References}
%%   \begin{small}
%%   \phantomsection
%%   \addcontentsline{toc}{section}{\bibname}
%%   \bibliographystyle{acm}
%%   \bibliography{bitcoin,tor}
%%   \end{small}

%% \end{frame}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: